---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
``` {r unzipfile}
if (!file.exists("activity.csv")) {
        unzip("activity.zip")
}
```
``` {r loadfile}
activity <- read.csv("activity.csv")
```
``` {r tidyverse, results = "hide"}
library(tidyverse)
```
```{r preprocessing}
str(activity)

activity$date <- as.Date(activity$date, tryFormats = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

``` {r histogram1}
h1 <- 
activity %>%
        select(- interval) %>%
        group_by(date) %>%
        summarise(steps = sum(steps, na.rm =T)) %>%
        ggplot() +
        geom_histogram(aes(steps),
                           fill = "steelblue",
                       position = "dodge")+
        theme_minimal() +
        labs(title = "Total steps per day",
             x = "Steps per Day"
        )
h1
```
``` {r SMeanMedian1}
sumsteps <-
activity%>%
        group_by(date) %>%
        summarise(steps = sum(steps, na.rm = T))

smean <- round(mean(sumsteps$steps), 2)
smedian <- median(sumsteps$steps)
```

Activity's daily Steps mean and median are the following **`r smean`** and **`r smedian`**, respectively. 

## What is the average daily activity pattern?

``` {r gganimate}
library(gganimate)
```
``` {r TimeSeries1, cache = T}
activity %>%
        group_by(interval) %>%
        summarise(steps = mean(steps, na.rm =T)) %>%
        ggplot() +
        geom_line(aes(x = interval,
                     y = steps),
                 stat = "identity") +
        theme_minimal() +
        labs(title = "Average Daily Activity Pattern",
             x = "Interval",
             y = "Average Number of Steps") +
        transition_reveal(along = interval)

```
``` {r maxactivity}
activity %>%
        group_by(interval) %>%
        summarise(steps = mean(steps, na.rm =T)) %>%
        filter(steps == max(steps))
```

## Imputing missing values

``` {r countingNAs}
sapply(activity, function(x) sum(is.na(x)))
```
``` {r imputeFunction}
mvimpute <- function(data=data) {
        for (i in 1:ncol(data)) {
                if (class(data[,i]) %in% c("integer", "numeric")) {
                        if (sum(is.na(data[,i]))) {
                                data[is.na(data[,i]), i] <- mean(data[,i], na.rm =T)
                        }
                }
        }
        data <- as_tibble(data)
        data$date <- as.Date(data$date, tryFormats = "%Y-%m-%d")
        data
}
```
``` {r mvimpute}
NoNA.actv <- mvimpute(activity)

head(NoNA.actv)
sapply(NoNA.actv, function(x) sum(is.na(x)))
```
``` {r histogram2}
h2 <-
NoNA.actv %>%
        select(- interval) %>%
        group_by(date) %>%
        summarise(steps = sum(steps, na.rm =T)) %>%
        ggplot() +
        geom_histogram(aes(steps),
                           fill = "steelblue",
                       position = "dodge")+
        theme_minimal() +
        labs(title = "Total steps per day (imputed data)",
             x = "Steps per Day"
        )
```
``` {r HistComparision}
library(patchwork)

h1 | h2
```
``` {r SMeanMedian2}
sumsteps2 <-
NoNA.actv%>%
        group_by(date) %>%
        summarise(steps = sum(steps, na.rm = T))

smean2 <- format(mean(sumsteps2$steps), scientific = F)
smedian2 <- format(median(sumsteps2$steps), scientific = F)
```

Activity's daily Steps mean and median (after imputing missing values) are the following **`r smean2`** and **`r smedian2`**, respectively. 

## Are there differences in activity patterns between weekdays and weekends?

``` {r weekdays}
activity$wday <-  weekdays(activity$date)

activity <- 
activity %>%
  select(steps, interval, date, wday) %>%
  mutate(week = case_when(
      (grepl("(segunda|terça|quarta|quinta|sexta)-feira", activity$wday) == T) ~ "weekday",
      (grepl("(sábado|domingo)", activity$wday) == T) ~ "weekend"
))
```
``` {r weekdayplot}
tsw1 <-
activity %>%
  group_by(week, interval) %>%
  summarise(steps = mean(steps, na.rm =T)) %>%
  filter(week == "weekday") %>%
  ggplot() +
  geom_line(aes(x = interval,
                y = steps),
                 color = "steelblue",  
                 stat = "identity") +
        theme_minimal() +
        labs(title = "Average Weekday Activity Pattern",
             x = "Interval",
             y = "Average Number of Steps") 
```
``` {r weekendplot}
tsw2 <-
activity %>%
  group_by(week, interval) %>%
  summarise(steps = mean(steps, na.rm =T)) %>%
  filter(week == "weekend") %>%
  ggplot() +
  geom_line(aes(x = interval,
                y = steps),
                 color = "steelblue",  
                 stat = "identity") +
        theme_minimal() +
        labs(title = "Average Weekend Activity Pattern",
             x = "Interval",
             y = "Average Number of Steps") 
```
``` {r pannelplot}
tsw1 / tsw2
```














